---
title: "Lab 08 - University of Edinburgh Art Collection"
author: "Barbara Mu"
date: ""
output: github_document
---

```{r setup, message = FALSE, warning = FALSE}
library(tidyverse)
library(skimr)
library(rvest)
```

## Load Packages and Data

First, let's load the necessary packages (done above in setup chunk).

Now, load the dataset:

```{r load-data, message = FALSE}
uoe_art <- read_csv("data/uoe-art.csv")
```

---

## Exercise 1

Fix the relative links by replacing the leading `"."` with the base URL:

```{r ex1-links}
# set url
first_url <- "https://collections.ed.ac.uk/art/search/*:*/Collection:%22edinburgh+college+of+art%7C%7C%7CEdinburgh+College+of+Art%22?offset=0"

# read html page
page <- read_html(first_url)

# scrape titles
titles <- page %>%
  html_nodes(".iteminfo") %>%
  html_node("h3 a") %>%
  html_text() %>%
  str_squish()

# scrape links and fix relative URLs
links <- page %>%
  html_nodes(".iteminfo") %>%
  html_node("h3 a") %>%
  html_attr("href") %>%
  str_replace("\\.", "https://collections.ed.ac.uk/art")

head(links)
```

The scraped `href` values are relative links (e.g., `./record/99803?...`), whereas the actual page URL is `https://collections.ed.ac.uk/art/record/99803?...`. We use `str_replace()` to swap the leading `.` with the full base URL.

## Exercise 2

Scrape artist names:

```{r ex2-artists}
artists <- page %>%
  html_nodes(".iteminfo") %>%
  html_node(".artist") %>%
  html_text() %>%
  str_squish()

artists
```

## Exercise 3

Put it all together in a tibble:

```{r ex3-tibble}
first_ten <- tibble(
  title  = titles,
  artist = artists,
  link   = links
)

first_ten
```

## Exercise 4

Scrape the second page:

```{r ex4-second-page}
second_url <- "https://collections.ed.ac.uk/art/search/*:*/Collection:%22edinburgh+college+of+art%7C%7C%7CEdinburgh+College+of+Art%22?offset=10"

page2 <- read_html(second_url)

titles2 <- page2 %>%
  html_nodes(".iteminfo") %>%
  html_node("h3 a") %>%
  html_text() %>%
  str_squish()

links2 <- page2 %>%
  html_nodes(".iteminfo") %>%
  html_node("h3 a") %>%
  html_attr("href") %>%
  str_replace("\\.", "https://collections.ed.ac.uk/art")

artists2 <- page2 %>%
  html_nodes(".iteminfo") %>%
  html_node(".artist") %>%
  html_text() %>%
  str_squish()

second_ten <- tibble(
  title  = titles2,
  artist = artists2,
  link   = links2
)

second_ten
```

## Exercise 5 & 6

Define the `scrape_page()` function and test it:

```{r ex5-function}
scrape_page <- function(url) {
  page <- read_html(url)

  titles <- page %>%
    html_nodes(".iteminfo") %>%
    html_node("h3 a") %>%
    html_text() %>%
    str_squish()

  links <- page %>%
    html_nodes(".iteminfo") %>%
    html_node("h3 a") %>%
    html_attr("href") %>%
    str_replace("\\.", "https://collections.ed.ac.uk/art")

  artists <- page %>%
    html_nodes(".iteminfo") %>%
    html_node(".artist") %>%
    html_text() %>%
    str_squish()

  tibble(
    title  = titles,
    artist = artists,
    link   = links
  )
}
```

```{r ex6-test, eval = FALSE}
scrape_page(first_url)
scrape_page(second_url)
```

## Exercise 7

Build the list of all URLs:

```{r ex7-urls}
root <- "https://collections.ed.ac.uk/art/search/*:*/Collection:%22edinburgh+college+of+art%7C%7C%7CEdinburgh+College+of+Art%22?offset="

urls <- paste0(root, seq(0, 3320, by = 10))

length(urls)
head(urls)
```

## Exercise 8

Map over all URLs to scrape all pages (pre-run; not evaluated here to save time):

```{r ex8-scrape, eval = FALSE}
# Takes several minutes to run
uoe_art_scraped <- map_dfr(urls, scrape_page)
```

## Exercise 9

Save the scraped data frame to the data folder (pre-run):

```{r ex9-write, eval = FALSE}
write_csv(uoe_art_scraped, "data/uoe-art.csv")
```

---

## Exercise 10

Read in the data and clean up the title column by separating the title and date:

```{r ex10-clean, warning = FALSE}
uoe_art <- read_csv("data/uoe-art.csv", show_col_types = FALSE)

uoe_art <- uoe_art %>%
  separate(title, into = c("title", "date"), sep = "\\(", extra = "merge", fill = "right") %>%
  mutate(
    year = str_remove(date, "\\).*") %>% as.numeric(),
    title = str_squish(title)
  ) %>%
  select(title, date, artist, year, link)
```

The warnings "NAs introduced by coercion" occur when `as.numeric()` encounters date strings that are not plain 4-digit years (e.g., `"1920s"`, `"19 Nov 1958"`, `"Unknown"`). This is expected and acceptable because our goal is only to capture year where it is a clean number.

## Exercise 11

```{r ex11-skim}
skim(uoe_art)
```

There are **108** works missing artist information and **1575** works missing year information.

## Exercise 12

```{r ex12-histogram}
ggplot(uoe_art, aes(x = year)) +
  geom_histogram(binwidth = 10) +
  labs(
    title = "Distribution of artwork years",
    x = "Year",
    y = "Count"
  )
```

Yes — there is one piece with a year of 2, which is clearly out of the ordinary for an art collection whose other works mostly date from the 1800s onward.

## Exercise 13

```{r ex13-outlier}
uoe_art %>%
  filter(!is.na(year)) %>%
  arrange(year) %>%
  select(title, artist, year) %>%
  head(5)
```

The piece **"Death Mask (2)"** by H. Dempshall has `year = 2`. The `(2)` in the title refers to a numbering convention (it is the 2nd piece in a series), not a date. Our code naively extracted `2` as the year because it splits on the first `(` and converts what follows to a number. The correct year, found on the art collection website, is **1964**.

```{r ex13-fix}
uoe_art <- uoe_art %>%
  mutate(year = if_else(title == "Death Mask", 1964, year))
```

```{r ex13-histogram-fixed}
ggplot(uoe_art, aes(x = year)) +
  geom_histogram(binwidth = 10) +
  labs(
    title = "Distribution of artwork years (corrected)",
    x = "Year",
    y = "Count"
  )
```

## Exercise 14

```{r ex14-artist}
uoe_art %>%
  count(artist, sort = TRUE) %>%
  filter(!is.na(artist)) %>%
  head(5)
```

The most commonly featured named artist is **Emma Gillies**, with 175 pieces. Emma Gillies (1898–1948) was a Scottish ceramicist who studied and taught at Edinburgh College of Art, so the university's collection naturally holds a large number of her works.

## Exercise 15

```{r ex15-child}
uoe_art %>%
  filter(str_detect(title, regex("child", ignore_case = TRUE))) %>%
  select(title, artist, year)
```

```{r ex15-count}
uoe_art %>%
  filter(str_detect(title, regex("child", ignore_case = TRUE))) %>%
  nrow()
```

There are **11** art pieces with the word "child" (or "Child") in their title. Using `ignore_case = TRUE` inside `regex()` captures both capitalised and lowercase forms.
