---
title: "Lab 08 - University of Edinburgh Art Collection"
author: "Barbara Mu"
date: ""
output: github_document
---

```{r}
setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) #set wrking directory to file location
direct = setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) #set wrking directory to file location
```

```{r}
library(robotstxt)
paths_allowed("https://collections.ed.ac.uk/art)")
```

## Load Packages and Data  

First, let's load the necessary packages:  

```{r load-packages, message = FALSE}
library(tidyverse) 
library(skimr)
library(rvest)
```

Now, load the dataset. If your data isn't ready yet, you can leave `eval = FALSE` for now and update it when needed.  

```{r}
# set url
first_url <- "https://collections.ed.ac.uk/art/search/*:*/Collection:%22edinburgh+college+of+art%7C%7C%7CEdinburgh+College+of+Art%22?offset=0"

# read html page
page <- read_html(first_url)
```

```{r}
titles <- page %>%
  html_nodes(".iteminfo") %>%
  html_node("h3 a") %>%
  html_text() %>%
  str_squish()
```

# Execrise 1 

```{r}
links <- page %>%
  html_nodes(".iteminfo") %>% 
  html_node("h3 a") %>% 
  html_attr("href") %>%
  str_replace("\\.", "https://collections.ed.ac.uk/art")

head(links)
```

# Execrise 2 

```{r}
artists <- page %>%
  html_nodes(".iteminfo") %>%
  html_node(".artist") %>%
  html_text() %>%
  str_squish()
```

# Execrise 3 

```{r}
first_ten <- tibble(
  title = titles,
  artist = artists,
  link = links)
```

# Execrise 4 

```{r}
second_url <- "https://collections.ed.ac.uk/art/search/*:*/Collection:%22edinburgh+college+of+art%7C%7C%7CEdinburgh+College+of+Art%22?offset=10"

page <- read_html(second_url)

titles <- page %>%
  html_nodes(".iteminfo") %>%
  html_node("h3 a") %>%
  html_text() %>%
  str_squish()

links <- page %>%
  html_nodes(".iteminfo") %>%
  html_node("h3 a") %>%
  html_attr("href") %>%
  str_replace("\\.", "https://collections.ed.ac.uk/art")

artists <- page %>%
  html_nodes(".iteminfo") %>%
  html_node(".artist") %>%
  html_text() %>%
  str_squish()

second_ten <- tibble(title = titles, artist = artists, link = links)
```

# Execrise 5 & 6

```{r}
source("scripts/02-scrape-page-function.R")
```

```{r}
scrape_page <- function(url) {
  page <- read_html(url)
  
  titles <- page %>%
    html_nodes(".iteminfo") %>%
    html_node("h3 a") %>%
    html_text() %>%
    str_squish()
  
  links <- page %>%
    html_nodes(".iteminfo") %>%
    html_node("h3 a") %>%
    html_attr("href") %>%
    str_replace("\\.", "https://collections.ed.ac.uk/art")
  
  artists <- page %>%
    html_nodes(".iteminfo") %>%
    html_node(".artist") %>%
    html_text() %>%
    str_squish()
  
  tibble(title = titles, artist = artists, link = links)
}

# Test it
scrape_page(first_url)
scrape_page(second_url)
```

# Execrise 7 
```{r}
root <- "https://collections.ed.ac.uk/art/search/*:*/Collection:%22edinburgh+college+of+art%7C%7C%7CEdinburgh+College+of+Art%22?offset="

urls <- paste0(root, seq(0, 3320, by = 10))
```

# Execrise 8 

```{r}
uoe_art <- map_dfr(urls, scrape_page)
```

# Execrise 9

```{r}
write_csv(uoe_art, "data/uoe-art.csv")
```

## Exercise 10
```{r}
uoe_art <- read_csv("data/uoe-art.csv")
```

```{r separate-title-date, error = TRUE}
uoe_art <- uoe_art %>%
  separate(title, into = c("title", "date"), sep = "\\(") %>%
  mutate(year = str_remove(date, "\\).*") %>% as.numeric()) %>%
  select(title, artist, year, link)
```

The warnings say "NAs introduced by coercion". This happens when as.numeric() is applied to date strings that aren't plain 4-digit years (e.g., "1920s", "19 Nov 1958", "Unknown"). This is expected and acceptable because our goal is only to capture clean year values where possible.

## Exercise 11  

```{r}
skim(uoe_art)
```

There are 108 works missing artist information and 1575 works missing year information. 

## Exercise 12  

```{r}
ggplot(uoe_art, aes(x = year)) +
  geom_histogram(binwidth = 10) +
  labs(title = "Distribution of artwork years",
       x = "Year", y = "Count")
```

## Exercise 13

```{r}
uoe_art %>%
  filter(!is.na(year)) %>%
  arrange(year) %>%
  select(title, artist, year) %>%
  head(5)
```

The piece Death Mask by H. Dempshall has year = 2, which is clearly an error. The `(2)` in the title refers to a numbering convention (it's the 2nd piece in a series), not a date. Our code naively extracted 2 as the year because it just looks for the first number inside parentheses. The correct year, found by visiting the piece's page on the art collection website, is 1964.

```{r fix-outlier}
uoe_art <- uoe_art %>%
  mutate(year = if_else(title == "Death Mask ", 1964, year))
```

```{r histogram-fixed}
ggplot(uoe_art, aes(x = year)) +
  geom_histogram(binwidth = 10) +
  labs(title = "Distribution of artwork years (corrected)",
       x = "Year", y = "Count")
```

## Exercise 14

```{r most-common-artist}
uoe_art %>%
  count(artist, sort = TRUE) %>%
  filter(!is.na(artist)) %>%
  head(5)
```

The most commonly featured artist (excluding unknowns) is Emma Gillies, with 175 pieces. Emma Gillies (1898–1948) was a Scottish ceramicist and ECA alumna. The University likely has so many of her pieces because she studied and taught at Edinburgh College of Art, making her work closely tied to the institution's history.

## Exercise 15

```{r child-in-title}
uoe_art %>%
  filter(str_detect(title, regex("child", ignore_case = TRUE))) %>%
  select(title, artist, year)
```

```{r child-count}
uoe_art %>%
  filter(str_detect(title, regex("child", ignore_case = TRUE))) %>%
  nrow()
```

There are 11 art pieces with the word "child" in their title.

